### 2018-06-12 Risk estimation/risk-estim.tex
Update ignore, begin df doc

\documentclass[11pt]{article}

% Include statements
\usepackage{graphicx}
\usepackage{amsfonts,amssymb,amsmath,amsthm}
\usepackage[numbers,square,sort]{natbib}
\usepackage[left=1in,top=1in,right=1in,bottom=1in,nohead]{geometry}
\usepackage{pdfsync}
\usepackage{multirow}
\usepackage{enumitem}
\setlist{nosep}
\usepackage{hyperref}
\hypersetup{backref,colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue}
\usepackage{algorithm,algorithmic}
% \SetKw{KwSet}{Set}
% \SetKw{KwSolve}{Solve}
\def\algorithmautorefname{Algorithm}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{6pt plus
  4pt minus 2pt}
\titlespacing\subsection{0pt}{6pt plus 4pt minus 2pt}{6pt plus
  4pt minus 2pt}
\titlespacing\paragraph{0pt}{6pt plus 4pt minus 2pt}{2pt plus
  2pt minus 0pt}

\usepackage[showonlyrefs]{mathtools}
%\usepackage{autonum}

% Bibliography

\renewcommand*{\figureautorefname}{Figure}%
\renewcommand*{\tableautorefname}{Table}%
\renewcommand*{\partautorefname}{Part}%
\renewcommand*{\chapterautorefname}{Chapter}%
\renewcommand*{\sectionautorefname}{Section}%
\renewcommand*{\subsectionautorefname}{Section}%
\renewcommand*{\subsubsectionautorefname}{Section}%

% Macros
\DeclareMathOperator*{\argmin}{argmin}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\one}{\mathbf{1}}
%\renewcommand{\S}{\mathcal{S}}
\newcommand{\Expect}[1]{\E\left[#1\right]}
\newcommand{\given}{\mbox{ }\vert\mbox{ }}
\renewcommand{\hat}{\widehat}
\DeclareMathOperator*{\trace}{tr}
\DeclareMathOperator*{\diag}{diag}
\renewcommand{\tilde}{\widetilde}
\newcommand{\Expectwrt}[2]{\mathbb{E}_{ #1 }\left[ #2 \right]}
\usepackage{xspace}
\makeatletter
\newcommand*{\iid}{%
    \@ifnextchar{.}%
        {i.i.d.}%
        {i.i.d.\@\xspace}%
}
\makeatother

\newcommand{\Pnd}{P_{\mathcal{N}(D)}}

\newtheorem{result}{Result}
\newtheorem{theorem}{Theorem}[result]
\renewcommand*{\sectionautorefname}{Section}%

\renewcommand{\P}{\mathbb{P}}
\DeclareMathOperator*{\argmax}{argmax}
\def\indep{\perp\!\!\!\perp}

\newcommand{\makeHeader}{\begin{center} 
DJM \hfill Risk estimation for Gen-Gen Lasso \hfill \today

\rule{\textwidth}{1pt}
\end{center}
}

\begin{document}
\makeHeader

This document combines a number of threads for model selection in our
problem (and other related optimization problems). First, we consider
estimating the degrees of freedom. Then we give two results which can
be used to select models.

\section{Degrees of freedom for Gen-Gen Lasso}
\label{sec:degrees-freedom-gen}

Suppose we are minimizing the negative log likelihood from a natural
exponential family subject to the generalized lasso penalty. That is,
given $Y_i \sim p(y \given \eta_i(\tau))$ for $\tau \in \R^p$, $i=1,\ldots,n$, our goal is to solve
\begin{equation}
  \label{eq:1}
  \hat\tau = \argmin_\tau \sum_{i=1}^n \phi(\eta_i(\tau)) - \langle y_i,\
  \eta_i(\tau)\rangle + \lambda\norm{D\tau}_1,
\end{equation}
where $D\in \R^{q\times p}$.

This generalizes the various standard models. For example,
$\ell_1$-trend filtering has $\phi(x) = x^2/2$, $\eta_i(x)=x_i$, and
$D$ the first-order discrete difference operator. Variance estimation
in our context has $\phi(x) = x$, $y_i = z^2_i$ (where $z_i\sim
N(0,\exp(-h)$) and $\eta_i(x) = -e^{-\tau_i}$. Logistic loss also
falls into this category.

We first consider the case that $\eta_i(\tau) = \eta_i(\tau_i)$, i.e.,
$\tau \in \R^n$ and $p(y_i \given \tau) = p(y_i \given \tau_i)$.

\begin{theorem}
  The divergence of $\hat\tau(y)$ is given by
  \begin{equation}
    \label{eq:2}
    \trace\left(D \hat\tau(y)\right) = -\Pnd \left(\Pnd
      \diag\left(\frac{d^2}{d\tau_i^2}\ell
        \vert_{y_i,\hat{\eta}_i}\right) \Pnd\right)^\dagger \Pnd
      \diag\left(\frac{d^2}{d\tau_i d y_i} \ell \vert_{y_i,\hat{\eta}_i}\right),
    \end{equation}
    where $\ell = \ell(y,\eta(\tau_i)) = \sum_{i=1}^n \phi(\eta_i(\tau_i)) - \langle y_i,\
    \eta_i(\tau_i)\rangle$, $\hat{\eta_i} = \eta_i(\hat\tau_i)$, and 
    \begin{equation}
      \label{eq:3}
      \Pnd = I_n - D_S^\top (D_S^\top D_S)^\dagger D_S
    \end{equation}
    is the projection onto the null-space of $D_S$, where $S = \{j \in
    [q] : D\hat\tau = 0\}$, and the notation $D_S$ means the rows of
    $D$ whose indices are in $S$.
\end{theorem}

\paragraph{Special cases:}
\begin{enumerate}
\item If we are interested in the natural exponential family (that is
  $\eta_i(\tau_i)=\tau_i$), then $\frac{d^2}{d\tau_i d y_i} \ell = -1$
  and $\frac{d^2}{d\tau_i^2}\phi_i(\eta_i(\tau_i)) = \textrm{Var}[y_i]$.
\item For Gaussian likelihood, $\frac{d^2}{d\tau_i^2}\ell = 1$
  and $\frac{d^2}{d\tau_i d y_i} \ell = -1$, so the divergence is the
  dimension of $\mathcal{N}(D_S)$ as shown in
  \citet{TibshiraniTaylor2012}. 
\item For logistic loss, $\frac{d^2}{d\tau_i^2}\ell =
  e^{\tau_i}/(1+e^{\tau_i})^2$ and 
  $\frac{d^2}{d\tau_i d y_i} \ell = -1$.
\item For the case of variance estimation, we have $\frac{d^2}{d\tau_i^2}\ell =
  y^2_ie^{-h}$ and 
  $\frac{d^2}{d\tau_i d y_i} \ell = -e^{-h}$.
\end{enumerate}

Now we generalize to the regression setting. Define
$\mu_i = x_i^\top \beta$. Furthermore, write
\begin{equation}
  \label{eq:4}
  \hat\beta = \argmin_\tau \sum_{i=1}^n \phi(\mu_i(\beta)) - \langle y_i,\
  \mu_i(\beta)\rangle + \lambda\norm{D\beta}_1.
\end{equation}

\begin{theorem}
  The divergence of $\hat\mu(y)$ is given by
  \begin{equation}
    \label{eq:5}
    \trace\left(D \hat\mu(y)\right) = -X_P \left(X^\top_P
      \diag\left(\frac{d^2}{d\mu_i^2}\ell
        \vert_{y_i,\hat{\mu}_i}\right) X_P\right)^\dagger X^\top_P
      \diag\left(\frac{d^2}{d\mu_i d y_i} \ell \vert_{y_i,\hat{\mu}_i}\right),
    \end{equation}
    where $\ell = \ell(y,\eta(\mu_i)) = \sum_{i=1}^n \phi(\eta_i(\mu_i)) - \langle y_i,\
    \mu_i\rangle$, $\hat{\eta_i} = \eta_i(x_i^\top\hat\beta)$, and 
    \begin{equation}
      \label{eq:6}
      X_P =( I_n - D_S^\top (D_S^\top D_S)^\dagger D_S) X
    \end{equation}
    is the projection of $X$ onto the null-space of $D_S$.
\end{theorem}

\begin{proof}
  This follows mainly from Theorem 2 in~\citep{VaiterDeledalle2017},
  though the conditions are non-trivial. To be expanded.
\end{proof}

\bibliographystyle{mybibsty}
\bibliography{AllReferences}
\end{document}