### 2018-06-29 Risk estimation/risk-estim.tex
Drafty but complete risk estimation

\documentclass[11pt]{article}

% Include statements
\usepackage{graphicx}
\usepackage{amsfonts,amssymb,amsmath,amsthm}
\usepackage[numbers,square,sort]{natbib}
\usepackage[left=1in,top=1in,right=1in,bottom=1in,nohead]{geometry}
\usepackage{pdfsync}
\usepackage{multirow}
\usepackage{enumitem}
\setlist{nosep}
\usepackage{hyperref}
\hypersetup{backref,colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue}
\usepackage{algorithm,algorithmic}
% \SetKw{KwSet}{Set}
% \SetKw{KwSolve}{Solve}
\def\algorithmautorefname{Algorithm}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{6pt plus
  4pt minus 2pt}
\titlespacing\subsection{0pt}{6pt plus 4pt minus 2pt}{6pt plus
  4pt minus 2pt}
\titlespacing\paragraph{0pt}{6pt plus 4pt minus 2pt}{2pt plus
  2pt minus 0pt}

%\usepackage{autonum}

% Bibliography

\renewcommand*{\figureautorefname}{Figure}%
\renewcommand*{\tableautorefname}{Table}%
\renewcommand*{\partautorefname}{Part}%
\renewcommand*{\chapterautorefname}{Chapter}%
\renewcommand*{\sectionautorefname}{Section}%
\renewcommand*{\subsectionautorefname}{Section}%
\renewcommand*{\subsubsectionautorefname}{Section}%

% Macros
\DeclareMathOperator*{\argmin}{argmin}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\one}{\mathbf{1}}
%\renewcommand{\S}{\mathcal{S}}
\newcommand{\Expect}[1]{\E\left[#1\right]}
\newcommand{\given}{\mbox{ }\vert\mbox{ }}
\renewcommand{\hat}{\widehat}
\DeclareMathOperator*{\trace}{tr}
\DeclareMathOperator*{\Cov}{Cov}
\DeclareMathOperator*{\diag}{diag}
\renewcommand{\tilde}{\widetilde}
\newcommand{\Expectwrt}[2]{\mathbb{E}_{ #1 }\left[ #2 \right]}
\usepackage{xspace}
\makeatletter
\newcommand*{\iid}{%
    \@ifnextchar{.}%
        {i.i.d.}%
        {i.i.d.\@\xspace}%
}
\makeatother

\newcommand{\Pnd}{P_{\mathcal{N}(G)}}

\usepackage{aliascnt}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newaliascnt{lemma}{theorem}
\newtheorem{lemma}[lemma]{Lemma}
\aliascntresetthe{lemma}
\providecommand*{\lemmaautorefname}{Lemma}
\newaliascnt{cor}{theorem}
\newtheorem{corollary}[cor]{Corollary}
\aliascntresetthe{cor}
\providecommand*{\corautorefname}{Corollary}
\newaliascnt{def}{theorem}
\newtheorem{definition}[def]{Definition}
\aliascntresetthe{def}
\providecommand*{\defautorefname}{Definition}
\newtheorem{example}{Example}

\renewcommand{\P}{\mathbb{P}}
\DeclareMathOperator*{\argmax}{argmax}
\def\indep{\perp\!\!\!\perp}

\usepackage{mathtools}
\mathtoolsset{showonlyrefs}

\newcommand{\makeHeader}{\begin{center} 
DJM \hfill Risk estimation for Gen-Gen Lasso \hfill \today

\rule{\textwidth}{1pt}
\end{center}
}

\begin{document}
\makeHeader

This document combines a number of threads for model selection in our
problem (and other related optimization problems). First, we consider
estimating the degrees of freedom. Then we give two results which can
be used to select models.

\section{Degrees of freedom for Gen-Gen Lasso}
\label{sec:degrees-freedom-gen}

Consider a random vector $Y \in \R^n$ with distribution in the
exponential family with independent
components. Specifically, write the density of $Y$ as
\begin{equation}
  \label{eq:4}
  p(Y \given \theta) = \left(\prod_{i=1}^n h(y_i)\right)\exp\left\{
    \sum_{i=1}^n w(y_i) \mu_i(\theta) -\psi_i(\theta) \right\}.
\end{equation}
We will consider estimation of $\theta$ subject to the generalized lasso penalty. That is,
our goal is to solve 
\begin{equation}
  \label{eq:1}
  \hat\theta(y) = \argmin_\theta \ell(y | \theta) +
  \lambda\norm{G\theta}_1 = \argmin_\theta\sum_{i=1}^n \psi_i(\theta) - w(y_i)\mu_i(\theta) + \lambda\norm{G\theta}_1,
\end{equation}
where $G\in \R^{q\times p}$ and $\theta\in\R^p$.

This generalizes many standard models. For example,
$\ell_1$-trend filtering has $p=n$, $\psi_i(x) = x_i^2/2$,
$\mu_i(x)=x_i$, $w(x)=2$, and
$G$ the first-order discrete difference operator. Variance estimation
in our context has $\psi_i(x) = x_i$, $w(x) = y^2_i$ and $\mu_i(x) = e^{-x_i}$. Logistic loss also
falls into this category. We use $G$ rather than the more common
choice of $D$ to avoid confusion with the differential operator that
we will require below.

\paragraph{Notation} For a vector-valued function $f:\R^p\rightarrow
\R^q$, we use the notation $Df$ to denote the Jacobian matrix
and $\nabla f$ to be the gradient.

We first consider the case that $\theta \in\R^n$ and that $\psi_i$ and
$\mu_i$ operate component-wise on
$\theta$. That is $p(y \given \theta) = \prod_{i=1}^n p(y_i \given \theta_i)$.

\begin{theorem}
\label{thm:simple-divergence}
  Suppose $p(y \given \theta) = \prod_{i=1}^n p(y_i \given \theta_i)$
  and that $\hat{\theta}$ is a solution of \eqref{eq:1}. Then, the
  divergence of $\hat\theta(y)$ is given by 
  \begin{equation}
    \label{eq:2}
    \trace\left(D \hat\theta(y)\right) = -\Pnd \left(\Pnd
      \diag\left(\frac{d^2}{d\theta_i^2}\ell
        \bigg\vert_{y_i,\hat{\theta}_i}\right) \Pnd\right)^\dagger \Pnd
      \diag\left(\frac{d^2}{d\theta_i d y_i} \ell \bigg\vert_{y_i,\hat{\theta}_i}\right),
    \end{equation}
    where 
    \begin{equation}
      \label{eq:3}
      \Pnd = I_n - G_S^\top (G_S^\top G_S)^\dagger G_S
    \end{equation}
    is the projection onto the null-space of $G_S$, where $S = \{j \in
    [q] : G\hat\theta = 0\}$, and the notation $G_S$ means the rows of
    $G$ whose indices are in $S$.
\end{theorem}

\paragraph{Special cases:}
\begin{enumerate}
\item If we are interested in the natural exponential family (that is
  $\mu_i(\theta)=\theta_i$ and $w(y_i)=y_i$), then $\frac{d^2}{d\theta_i d y_i} \ell = -1$
  and $\frac{d^2}{d\theta_i^2}\psi_i(\theta_i) = \textrm{Var}[Y_i]$.
\item In particular, if $p(y_i \given \theta_i) = \mbox{N}(\theta_i, \sigma^2)$, then $\frac{d^2}{d\theta_i^2}\ell = \sigma^2$
  and $\frac{d^2}{d\tau_i d y_i} \ell = -1$, so the divergence is
  $\sigma^2$ times the
  dimension of $\mathcal{N}(G_S)$ as shown in
  \citet{TibshiraniTaylor2012}. 
\item For logistic loss, $\frac{d^2}{d\theta_i^2}\ell =
  e^{\theta_i}/(1+e^{\theta_i})^2$ and 
  $\frac{d^2}{d\tau_i d y_i} \ell = -1$.
\item For the case of variance estimation, we have $\frac{d^2}{d\theta_i^2}\ell =
  y^2_ie^{-\theta}$ and 
  $\frac{d^2}{d\theta_i d y_i} \ell = -e^{-\theta}$.
\end{enumerate}

Now we generalize to the regression setting. Define
$\theta_i = x_i^\top \beta$, $\beta \in \R^p$. And let
\begin{equation}
  \label{eq:7}
   \hat\beta(y) = \argmin_\beta \sum_{i=1}^n \psi_i(x_i^\top \beta) -
   w(y_i)\mu_i(x_i^\top \beta) + \lambda\norm{G\beta}_1,
\end{equation}

\begin{theorem}
\label{thm:regression-divergence}
  The divergence of $\hat\theta(y) := X\hat\beta(y)$ is given by
  \begin{equation}
    \label{eq:5}
    \trace\left(D \hat\theta(y)\right) = -X_P \left(X^\top_P
      \diag\left(\frac{d^2}{d\theta_i^2}\ell
        \bigg\vert_{y_i,\hat{\theta}_i}\right) X_P\right)^\dagger X^\top_P
      \diag\left(\frac{d^2}{d\theta_i d y_i} \ell \bigg\vert_{y_i,\hat{\theta}_i}\right),
    \end{equation}
    where 
    \begin{equation}
      \label{eq:6}
      X_P =( I_n - G_S^\top (G_S^\top G_S)^\dagger G_S) X
    \end{equation}
    is the projection of $X$ onto the null-space of $G_S$.
\end{theorem}

\begin{proof}
  This follows mainly from Theorem 2 in~\citep{VaiterDeledalle2017},
  though the conditions are non-trivial. To be expanded.
\end{proof}

\section{Risk estimation}
\label{sec:risk-estimation}

If $Y \sim \mbox{N}(\theta, \sigma^2)$, a now common method of risk
estimation makes use of Stein's Lemma.
\begin{lemma}[Stein's Lemma]
  Assume $f(Y)$ is weakly differentiable with essentially
  bounded weak partial derivatives on $\R^n$, then
  \begin{equation}
    \label{eq:8}
    \trace \Cov(Y,f(Y)) = \Expect{\left\langle Y,\ f(Y)\right\rangle} = \sigma^2\Expect{\trace Df(Y) \bigg\vert_y }.
  \end{equation}
\end{lemma}
The utility of this result comes from examining the decomposition of
the mean squared error of $f(Y)$ as an estimator of $\theta$.
\begin{align}
  \label{eq:9}
  \Expect{\norm{\theta-f(Y)}_2^2} 
  &= \Expect{\norm{Y-f(Y)}} -n\sigma^2 + 2
    \trace\Cov(Y,f(Y))\\
  &= \Expect{\norm{Y-f(Y)}} -n\sigma^2 + 2\sigma^2
    \Expect{\trace Df(y) \bigg\vert_Y} .
\end{align}
This characterization motivates the definition of degrees-of-freedom
for linear predictors ($\textrm{df} :=\frac{1}{\sigma^2} \trace
Df(y)\big\vert_Y$)~\citep{efron1986biased}, where $f(y)=Hy$. Using 
Stein's Lemma, assuming $\sigma^2$ is known, we have Stein's Unbiased
Risk Estimator
\begin{equation}
  \label{eq:10}
  SURE_\theta = \norm{Y-f(Y)} -n\sigma^2 + 2\sigma^2\trace Df(y) \bigg\vert_y.
\end{equation}
Note that this is a risk for estimating the $n$-dimensional parameter $\theta$.

The following result generalizes this idea to certain continuous
exponential families.
\begin{lemma}[Generalized Stein Lemma~\citep{Eldar2009}]
  \label{thm:g-stein-lemma}
  Assume $f(Y)$ is weakly differentiable with essentially
  bounded weak partial derivatives on $\R^n$. Let $Y$ be distributed
  according to a natural exponential family
  \begin{equation}
    \label{eq:11}
    p(Y \given \beta) = \left(\prod_{i=1}^n h(y_i)\right)\exp\left\{
    \sum_{i=1}^n y_i \theta_i(\beta_i) -\psi(\beta_i) \right\},
  \end{equation}
  and assume that $h$ is weakly differentiable and that $\theta_i$ is
  one-to-one. Then,
  \begin{equation}
    \label{eq:12}
    \Expect{\left\langle \theta(\beta),\ f(Y)\right\rangle} = - \Expect{
      \left\langle\frac{\nabla h(Y)}{h(Y)},\ f(Y)\right\rangle + \trace Df(y) \bigg\vert_Y}.
  \end{equation}
Note that $\nabla h(Y)$ here means the vector $[d/dy h(y)
\vert_{y_i}]$ and $h(Y)$ means the vector $[h(y_i)]$.
\end{lemma}
Therefore we define the Generalized SURE~\citep{Eldar2009} along the lines of the multivariate Gaussian case.
\begin{lemma}
  \label{lem:gsure}
  Assume $h$ and $\nabla h$ are weakly differentiable, $\theta_i$ is one-to-one,
  $f(Y)$ is weakly differentiable with essentially bounded partial
  derivatives and that $p(Y\given\beta)$ is as in \eqref{eq:11}. Then
  \begin{equation}
    \label{eq:13}
    GSURE_{\theta} = \norm{f(Y)}^2_2 + 2 \left\langle\frac{\nabla
        h(Y)}{h(Y)},\ f(Y)\right\rangle + 2\trace Df(y) \bigg\vert_Y +
    \frac{1}{h(Y)} \trace \frac{\partial^2 h(y)}{\partial y^2}\bigg\vert_Y
  \end{equation}
  is an unbiased estimator for the MSE of an estimator $f(Y)$ of
  $\theta$: $\Expect{\norm{f(Y) - \theta(\beta)}_2^2}$.
\end{lemma}

\begin{proof}
  We have
  \begin{align}
    \label{eq:14}
    \Expect{\norm{f(Y) - \theta(\beta)}_2^2} 
    &= \Expect{\norm{f(Y)}_2^2} + \Expect{\norm{\theta}_2^2} -
      2\Expect{\left\langle \theta(\beta),\ f(Y)\right\rangle}.
  \end{align}
  Now, the first term is a function of the data only, and to the last term, we simply apply
  \autoref{thm:g-stein-lemma}. For the second term,
  \begin{align}
    \Expect{\norm{\theta}_2^2} 
    &=\Expect{\langle \theta,\ \theta \rangle}
     = -\Expect{\left\langle\frac{\nabla h(Y)}{h(Y)},\ \theta\right\rangle}\\ 
    &= \Expect{\left\langle\frac{\nabla h(Y)}{h(Y)},\
      \frac{\nabla h(Y)}{h(Y)}\right\rangle} + \Expect{\trace
      \frac{\partial}{\partial y} \frac{\nabla h(y)}{h(y)}
      \bigg\vert_Y}\\
    &= \Expect{\frac{\norm{\nabla h(Y)}_2^2}{h(Y)^2}} + \Expect{\trace
      \frac{\norm{\nabla h(Y)}_2^2 +h(Y) \partial^2/\partial y^2 h(y)\big\vert_Y} 
      {h(Y)^2}}\\ 
    &=\Expect{ \frac{1}{h(Y)} \trace \frac{\partial^2 h(y)}{\partial y^2}\bigg\vert_Y},
  \end{align}
  by applying \autoref{thm:g-stein-lemma} twice along with the
  quotient rule.
\end{proof}

\begin{theorem}
  Let $Y_i \sim N(0, e^{x_i})$. Consider estimating $x$ by solving
  \begin{equation}
    \label{eq:15}
    \min_h \sum_{i=1}^n x_i + y_i^2e^{-x_i} + \lambda \norm{Gx}_1.
  \end{equation}
  Then an unbiased estimator of $\Expect{\norm{\frac{1}{2e^x} -
      \frac{1}{2e^{\hat x}}}}_2^2$ is given by
  \begin{equation}
    \label{eq:16}
    \norm{\frac{1}{2e^{\hat x}}}_2^2 - \frac{1}{2}\left\langle
      \frac{1}{y^2},\ \frac{1}{e^{\hat x}}\right\rangle + 2 \Pnd \left(\Pnd
      \diag\left(y^2e^{-\hat x}\right) \Pnd\right)^\dagger \Pnd
      \diag\left( e^{-\hat x}\right) + \frac{3}{4y^4},
  \end{equation}
where $\Pnd$ is as in \eqref{eq:3}.
\end{theorem}

\begin{proof}
  Define $z_i := y_i^2$ and $\theta_i = -\frac{1}{2e^{x_i}}$. Then
  \begin{align}
    \label{eq:17}
    p(z_i \given x_i) 
    &= \frac{1}{\sqrt{2\pi z_i e^{x_i}}} \exp\left\{-\frac{1}{2}
      \frac{z_i}{e^{x_i}}\right\}\one_{(0,\infty)}(z_i)\\
    &= \frac{1}{\sqrt{\pi z}}\one_{(0,\infty)}(z_i) \exp\left\{z_i \cdot
      \left(-\frac{1}{2e^{x_i}}\right) -
      \left(-\frac{1}{2}\log\left(\frac{1}{2 e^{x_i}}\right)
      \right)\right\}\\
    &= \frac{1}{\sqrt{\pi z}}\one_{(0,\infty)}(z_i)
      \exp\left\{z_i\cdot \theta_i - \psi(\theta_i)\right\},
  \end{align}
  which is a natural exponential family with $\psi(t) =
  -\frac{1}{2}\log(-t)$ and $h(t) = \frac{1}{\sqrt{\pi t}}
  \one_{(0,\infty)}(t)$. Therefore, $d/dt \log h(t) = -1/(2t)$ and
  $(d^2/dt^2 h(t))/h(t) = 3 / (4t^2)$. Finally, we have $\frac{d^2}{d\theta_i^2}\ell =
  y^2_ie^{-\theta}$ and 
  $\frac{d^2}{d\theta_i d y_i} \ell = -e^{-\theta}$. The result
  follows from \autoref{thm:simple-divergence} and \autoref{lem:gsure}.

\end{proof}

An alternate model selection technique is to examine the
Kullback-Leibler Divergence between the density under
$\theta=\hat\theta(Y)$ and that under $\theta=\theta$:
$\Expect{KL(p(Y\given \hat\theta(Y)) \Vert p(Y\given \theta))}$. For
exponential families, we have
\begin{equation}
\Expect{KL(\hat\theta(Y) \Vert 
  \theta)} = \Expect{KL(p(Y\given \hat\theta(Y)) \Vert p(Y\given
  \theta))} = \left\langle \hat\theta(Y)-\theta,\ \nabla
  \psi(\hat\theta(Y))\right\rangle + \psi(\theta) - \psi(\hat\theta(Y)).
\end{equation}

An application of \autoref{thm:g-stein-lemma} thus provides an
unbiased estimator of this quantity~\citep{Deledalle2017}.
\begin{lemma} 
  \label{lem:sukls}
  Assume $h$ and $\nabla h$ are weakly differentiable, $\theta_i$ is one-to-one,
  $f(Y)$ is weakly differentiable with essentially bounded partial
  derivatives and that $p(Y\given\beta)$ is as in \eqref{eq:11}. Then
  \begin{equation}
    SUKLS = \left\langle \hat\theta + \frac{\nabla h(Y)}{h(Y)},\
      \nabla \psi(\hat\theta(Y))\right\rangle + \trace Df(y) \bigg\vert_Y
    - \psi(\hat\theta(Y))
  \end{equation}
  is unbiased for $\Expect{KL(\hat\theta(Y) \Vert
  \theta)} - \psi(\theta)$.
\end{lemma}

\begin{corollary}
   Let $Y_i \sim N(0, e^{x_i})$. Consider estimating $x$ by solving
  \begin{equation}
    \label{eq:15}
    \min_h \sum_{i=1}^n x_i + y_i^2e^{-x_i} + \lambda \norm{Gx}_1.
  \end{equation}
  Then an unbiased estimator of $\Expect{KL(\hat x \Vert
    x} - \psi(\theta)$ is given by
  \begin{equation}
    \label{eq:18}
    -\frac{n}{2} - \left\langle \frac{1}{2y^2},\ e^{\hat
        x}\right\rangle + 
    -\frac{1}{2}\log\left(2e^{\hat
        x_i}\right) + \Pnd \left(\Pnd
      \diag\left(y^2e^{-\hat x}\right) \Pnd\right)^\dagger \Pnd
      \diag\left( e^{-\hat x}\right). 
  \end{equation}
\end{corollary}
\begin{proof}
  As $\hat \theta_i = -\frac{1}{2e^{\hat x_i}}$, 
  \[
    \psi(\hat\theta_i) = -\frac{1}{2}\log\left(\frac{1}{2e^{\hat
          x_i}}\right)
  \]
  and $d/d\theta \psi(\theta) = -\frac{1}{2\theta}$. Thus
  \begin{align}
    SUKLS 
    &= \left\langle -\frac{1}{2e^{\hat x}} - \frac{1}{2y^2},\ e^{\hat
      x}\right\rangle + \trace Df(y) \bigg\vert_Y
      +\frac{1}{2}\log\left(\frac{1}{2e^{\hat
      x_i}}\right)\\
    &= -\frac{n}{2} - \left\langle \frac{1}{2y^2},\ e^{\hat
      x}\right\rangle + \trace Df(y) \bigg\vert_Y
      -\frac{1}{2}\log\left(2e^{\hat
      x_i}\right).
  \end{align}
NOTE: I think some signs are wrong here.
\end{proof}

\bibliographystyle{mybibsty}
\bibliography{AllReferences}
\end{document}